\documentclass[fontsize=12pt, paper=a4]{scrlttr2}
\usepackage[english]{babel}
\usepackage[parfill]{parskip}  
\usepackage{geometry}
\usepackage{color} \definecolor{dark-gray}{gray}{0.3}
\usepackage[colorlinks=true,urlcolor=dark-gray,breaklinks=true,citecolor=black,linkcolor=black]{hyperref}
\usepackage{breakurl}

\setkomavar{fromname}{Will Pearse}
\setkomavar{fromaddress}{Department of Biology, McGill University\\1205 ave Docteur Penfield, Room W3/4\\Montreal, Quebec, H3A 1B1\\Canada}

\setkomavar{signature}{Will Pearse}
\renewcommand{\raggedsignature}{\raggedright} % make the signature ragged right
\makeatletter
\@setplength{refvpos}{\useplength{toaddrvpos}}
\makeatother

\setkomavar{subject}{Submission of ``\emph{pez}: Phylogenetics for the Environmental Sciences}
\KOMAoptions{fromalign=left, fromlogo=false, addrfield=no,
backaddress=no,  subject=left,foldmarks=no,
fromphone=no,fromemail=true,fromalign=right,pagenumber=no}
\setkomavar{fromemail}{will.pearse@gmail.com}
\begin{document}
\begin{letter}

\opening{To whom it may concern,}\enlargethispage{8\baselineskip}

I am writing to ask you to consider the manuscript `\emph{pez:
  Phylogenetics for the Environmental Sciences}', for publication in
\emph{Bioinformatics}. The manuscript describes an \emph{R} package
(\emph{pez}) that integrates and facilitates the analysis and
simulation of evolutionary and ecological data.

%Needs more reference to statistical models

An ever-increasing number of ecologists test hypotheses using
phylogenetic data, but there is no unified data structure or interface
that integrates ecological and phylogenetic data-types. Moreover,
there are a number of statistical methods that, despite their
popularity, have not been formally released as software packages and
as such are patchily and inconsistently implemented.

\emph{pez} solves the first of these problems and goes some way to
addressing the second. It integrates existing phylogenetic metrics
within a simple framework, and provides novel code for many metrics
that were previously unavailable. \emph{pez} also contains many new
modelling (\emph{e.g.}, Phylogenetic Generalised Linear Mixed Models;
PGLMMs) and simulation functions. It is my hope that \emph{pez} will
more tightly integrate the computational development of the field of
eco-phylogenetics, ensuring the fragmentation I believe \emph{pez} to
have reversed will not recur.

I feel this manuscript is best-suited to \emph{Bioinformatics} because
many conceptually related \emph{R} packages (\emph{e.g.},
\emph{picante} and \emph{geiger}) have been published there. The
package has been developed using modern, open software development
techniques, which I feel your journal is well-placed to review. I have
indicated potential reviewers in the online submisison who I feel have
implemented statistical methods similar to those in \emph{pez}.

\closing{Thank you for your time, and I look forward to hearing from
  you,\vspace{20pt}}

\end{letter}
\end{document}